Choose a License

### 2025-01-17 choose_license.qmd
Shorten section on other restrictions
- Make the argument more succinct
- Remove the references as they are
  not that good

As indicated before, a free/open license must allow creating derivative works and must allow exercising the rights granted by it _for any purpose_, including commercial use. There are a few good resources on the reasoning behind that [e.g., see @Klimpel2013; @ODI2015; @Stallman2022NoLimit; @Moller2023; @Vezina2020; @WikimediaCommons2009], but we would like to highlight one reason in particular.

With the rights provided by free/open licenses comes the possibility to build on and combine multiple works by different authors, which is essential for any cultural and scientific activity. However, this is only possible if the various licenses involved are compatible with each other. For example, one is allowed to remix two figures if the first is licensed under CC\ BY\ 4.0 and the second under CC\ BY-SA\ 4.0 because the licenses were written to be compatible with each other [compare @CC2024FAQ]. Also note that CC\ BY-SA\ 4.0 is [one-way compatible](https://creativecommons.org/compatiblelicenses) with the GPLv3, which in turn is compatible with the AGPLv3.

However, applying a restriction such as only permitting non-commercial use or not allowing derivative works creates _silos_ of works which are mutually incompatible with each other. Put differently, one cannot share a remix of two works where one is licensed under CC\ BY-SA\ 4.0 and the other only allows non-commercial use. In order to avoid silos, one should only choose licenses which are compatible with the dominant copyleft license in the respective community [@Lammerhirt2017; @Wheeler2014]. If you would like to learn more about the different types of compatibility, we recommend you to read the article "A Quick Guide to Software Licensing for the Scientist-Programmer" by @Morin2012. The following diagram provides an overview of the compatibility of various licenses:

>>>>

As indicated before, a free/open license must allow creating derivative works and must allow exercising the rights granted by it _for any purpose_, including commercial use. Otherwise, creating adaptations from multiple works by different authors would either limit commercial use for everybody (including the authors) or would not be possible at all (when including materials with a copyleft license), effectively creating _silos_ of works which are mutually incompatible with each other. This is especially relevant for work types for which adaptations are created frequently, such as text, data, and source code. In order to avoid silos, one should only choose licenses which are compatible with the dominant copyleft license in the respective community [@Lammerhirt2017; @Wheeler2014]. If you would like to learn more about the different types of compatibility, we recommend you to read the article "A Quick Guide to Software Licensing for the Scientist-Programmer" by @Morin2012. The following diagram provides an overview of the compatibility of various licenses:
